Wait-free Queues with Polygarithmic Step Complexity

### 2022-07-04 article/psuedocode.tex
add analysis

% \item In a propagate step at most 2 different time values are read \\ If there are more than 2 numbers then the smallest number should have been propagated far before.
% \item There are at most $p^2$ blocks with same time value in a node. \\ At most p processes could die before line 27 and each contains at most p elements.
\begin{lemma} \nf{super[i+1]-super[i]}$\leq p$\end{lemma}
\begin{proof}
 In a Refresh with successful CAS in line 46, \nf{super} and \nf{counter} are set for each child in lines 48,49. Assume the current value of the counter in node \nf{n} is \nf{i+1} and still \nf{super[i+1]} is not set. If an instance of successful \nf{Refresh(n)} finishes \nf{super[i+1]} is set a new value and a block is added after \nf{n.parent[sup[i]]}. There could be at most $p$ successful unfinished concurrent instances of \nf{Refresh()} that have not reached line 49. So the distance between \nf{super[i+1]} and \nf{super[i]} is less than $p$.
 \begin{figure}[hbt]
  \center\includegraphics[width=4in]{pics/index2}
\end{figure}  
\end{proof}
\begin{lemma}[super property]\label{superCounter}
If \nf{super[i] $\neq$ null} in node \nf{n}, then \nf{super[i]} is the index of the superblock of a block with \nf{time=i} in \nf{n.parent.blocks}.
\end{lemma}
\begin{lemma} Superblock of \nf{b} is within range $\pm 2p$ of the \nf{super[b.time]}.
\end{lemma}

@@

\begin{lemma}[Amortized time analysis]
%  \nf{n.GetEnq(b,i), n.Index(b,i)} take $O(\log^2 p)$ steps. Search in the root may take $O(\log Q+ p^2)$ steps. Helping is done every $p^2$ block appended to the root and takes $p\times \log^2p$ steps. Amortized time consumed for helping by each process is $O(\log^2 p)$.
\nf{Enqueue()} and \nf{Dequeue} take $O(\log62 p +q)$ steps (amortized anlysis), which $p$ is the number of processes and $q$ is the size of the queue at the time of invocation.
\end{lemma}

>>>>

% \item In a propagate step at most 2 different time values are read \\ If there are more than 2 numbers then the smallest number should have been propagated far before.
% \item There are at most $p^2$ blocks with same time value in a node. \\ At most p processes could die before line 27 and each contains at most p elements.
\begin{lemma} \nf{super[i+1]-super[i]}$\leq p$\end{lemma}
\begin{proof}
 In a Refresh with successful CAS in line 46, \nf{super} and \nf{counter} are set for each child in lines 48,49. Assume the current value of the counter in node \nf{n} is \nf{i+1} and still \nf{super[i+1]} is not set. If an instance of successful \nf{Refresh(n)} finishes \nf{super[i+1]} is set a new value and a block is added after \nf{n.parent[sup[i]]}. There could be at most $p$ successful unfinished concurrent instances of \nf{Refresh()} that have not reached line 49. So the distance between \nf{super[i+1]} and \nf{super[i]} is less than $p$.
 \begin{figure}[hbt]
  \center\includegraphics[width=4in]{pics/index2}
\end{figure}  
\end{proof}
\begin{lemma}[super property]\label{superCounter}
If \nf{super[i] $\neq$ null} in node \nf{n}, then \nf{super[i]} is the index of the superblock of a block with \nf{time=i} in \nf{n.parent.blocks}.
\end{lemma}
\begin{lemma}\label{superRange} Superblock of \nf{b} is within range $\pm 2p$ of the \nf{super[b.time]}.
\end{lemma}

@@

\begin{lemma}[Amortized time analysis]
%  \nf{n.GetEnq(b,i), n.Index(b,i)} take $O(\log^2 p)$ steps. Search in the root may take $O(\log Q+ p^2)$ steps. Helping is done every $p^2$ block appended to the root and takes $p\times \log^2p$ steps. Amortized time consumed for helping by each process is $O(\log^2 p)$.
\nf{Enqueue()} and \nf{Dequeue} take $O(\log^2 p +q)$ steps (amortized anlysis), which $p$ is the number of processes and $q$ is the size of the queue at the time of invocation.
\end{lemma}
\begin{proof}
\nf{Enqueue(x)} consists of creating a \nf{block(x)} and appending it to the tree. The first part takes constant time. To propagate \nf{x} to the root the algorithm tries two \nf{Refresh}es in each node of the path from the leaf to the root (Lines \ref{firstRefresh}, \ref{secondRefresh}). Each \nf{Refresh} takes $O(1)$ steps since creating a block is done in constant time and does $O(1)$ \nf{CAS}es. Since the height of the tree is $O(\log p)$, \nf{Enqueue(x)} takes $O(\log p)$ steps.

A \nf{Dequeue()} creates a block with null value element, appends it to the tree, computes its order among operations, and returns the response. The first two part is similar to an \nf{Enqueue} operation. To compute the order there are some constant steps and \nf{IndexDeq} is called. \nf{IndexDeq} does a search with range $p$ in each level (Lemma \ref{superRange}) which takes $O(log^2 p)$ in the tree. In the \nf{FindResponse()} routine \nf{DSearch()} in the root takes $\Theta(\log$(\nf{root.blocks[b].\size+root.blocks[end].\size}) by Lemma \ref{dsearch}, which is $O(\log$ size of the queue when \nf{enqueue} is invoked$)+\log$ size of the queue when \nf{dequeue} is invoked$)$. Each search in \nf{GetEnq()} takes $O(\log p)$ since there are $\leq p$ subblocks in a block (Lemma \ref{subBlocksBound}), so \nf{GetEnq()} takes $O(\log^2 p)$ steps.

If we split \nf{DSearch} time cost between the corresponding \nf{Enqueue}, \nf{Dequeue}, in amortized we have \nf{Enqueue} takes $O(\log p +q)$ and \nf{Dequeue} takes $O(\log^2 p +q)$ steps.
\end{proof}
